\papertitle

### 2020-01-23 paper/paper (copy).tex
Paper - First version

Assume, now, we have the scenario introduced in Figure \ref{fig:cylc-io-model-2}. In this case, we have extra information indicating why Tasks \texttt{ver} and \texttt{post} depend on Task \texttt{model}. In fact, the dependency is not to the complete task, but of some datasets produced by Task \texttt{model}. If this information is available a priori, then we can use it to start Task \texttt{ver} once we have the dataset \texttt{I/O 1} and start Task \texttt{post} once we have the dataset \texttt{I/O 3}, making the general process faster. Still considering this small example, we can assume that dataset \texttt{I/O 2} will not be used for other tasks. So this data can be stored in a "slow" storage or iot can even be deleted.

@@

>>>>

Assume, now, we have the scenario introduced in Figure \ref{fig:cylc-io-model-2}. In this case, we have extra information indicating why Tasks \texttt{ver} and \texttt{post} depend on Task \texttt{model}. In fact, the dependency is not to the complete task, but of some datasets produced by Task \texttt{model}. If this information is available a priori, then we can use it to start Task \texttt{ver} once we have the dataset \texttt{I/O 1} and start Task \texttt{post} once we have the dataset \texttt{I/O 3}, making the general process faster. Still considering this small example, we can assume that dataset \texttt{I/O 2} will not be used for other tasks. So this data can be stored in a "slow" storage or it can even be deleted.

@@

\section{Nice Pictures -- Version 2}

Consider the Cylc suite configuration for a toy monthly cycling workflow in Figure \ref{fig:cylc}. A warm-cycled atmospheric model (\textbf{model}) is followed by postprocessing (\textbf{post}), forecast verification (\textbf{ver}), and product generation (\textbf{prod}) tasks. Also, consider, in addition, a task \textbf{check} that compares some verification metric against products from two cycles earlier.

The first cycle of the suite in Figure \ref{fig:cylc} is again represented in Figure \ref{fig:cycle1}. This suite presents the dependencies among the tasks, but it is missing the information about I/O and storage. The decision about where the data with be stored is usually made by the developers in stages prior to Cylc and then it is incorporated into the Cylc process. The decisions about storage have to be made taking into consideration the architecture in which the workflow will run. Applied scientists should not spent time understanding hardware characteristics, but using their time to develop their work, let the computer do the job the simplest way possible, and just collect and analyse the results. However, nowadays, to be able to run a job in a HPC environment, researchers have to have a profound knowledge about their workflow, which is expected, but also about decisions regarding storage, communication, and computing. In this paper, we propose an approach to reduce the burden on researchers and, at the same time, optimise the decisions about jobs running in HPC systems. Once we have an automated decision about where the job will run and how the storage will be managed, the scientists can the run their workflows in any machine without further interactions and even without previous knowledge about the machine architecture.

Consider one cycle of the workflow, as shown in Figure \ref{fig:model7}. The tasks dependencies are represented by the arrows, but there is no information about the data that are generated. Tasks \texttt{ver} and \texttt{post} depend on Task \texttt{model}, but does Task \texttt{model} needs to be fully completed to satisfy the dependency? Of course once Task \texttt{model} is completed, then Tasks \texttt{ver} and \texttt{post} can proceed, and that is the information Cylc has to create and reproduce the workflow.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.4\columnwidth]{cycle1}
  \caption{First stage workflow}
  \label{fig:cycle1}
\end{figure}

The scenario prtesented in Figure \ref{fig:cycle-io} shows the same cycle of Figure \ref{fig:cycle-1} but now information about datasets is introduced. The way data moves through tasks is by creating datasets to pass on the information, including filename and storage destination. The datasets produced by every task are coded in the workflow scripts prior to Cylc evaluation and it is just replicated by this tool. The tasks can create a different number of datasets, and these datasets may be necessary for the immediate next task in the current cycle, or for Tasks in different cycles. Figure \ref{fig:cycle-io-dep} shows the dependencies among the datasets and the Tasks: datasets colours indicate with task will use the information and datasets in yellow represent information that will be used in a different cycle. Note that the information that is currently being used in the current cycle can also be used later.
Also note that one dataset produced by Task \texttt{model} is later used in Task \texttt{prod gen}. This is not an immediate conclusion when we analyse just the arrows for the dependencies among Tasks. It is clear that Task \texttt{model} is a dependency for Task \texttt{post prod} which is a dependency for Task \texttt{prod gen}. The point here is that the knowledge that, for this example, Task \texttt{model} will generate, one dataset for Task \texttt{post proc}, one dataset for Task \texttt{prod gen}, two datasets for Task \texttt{verification}, and three dataset for a later cycle can be used to optimise how these datasets are stored during the workflow. This information today is static and manual, leaving opened many opportunities for automatisations and optimisations.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{cycle-io}
  \caption{First stage workflow including datasets}
  \label{fig:cycle-io}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{cycle-io-dep}
  \caption{First stage workflow including datasets and dependencies}
  \label{fig:cycle-io-dep}
\end{figure}

Figure \ref{fig:cycle-4} exemplifies another restriction of the way current workflows are being generated. Assume you have, for instance, four cycles of the model are running. Because the scripts for the workflow consider the individual tasks and the datasets generated by them are placed at the same place at storage every time.

all datasets generated by Task \texttt{model} that will be used by Task \texttt{post prod}

all datasets generated by Task \texttt{A} and used by Task \texttt{B} will have the same physical localion in the system, because there is one fixed script responsible for setting the configuration. Again, such configuration was done manually and with restricted information about the storage and system architecture. It would be interesting to explore the options for the datasets and have them placed in different storage systems. If the lifetime of such files is, for instance, two cycles, alternate the files into two schatch file systems is something would be simple job for a tool working on generating filenames and storage. However, manually that implies at having at least two scripts with information about the different storage and manual work to have a better utilisation of the available resources.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{cycle-4}
  \caption{Four stages workflow example}
  \label{fig:cycle-4}
\end{figure}

Figure \ref{fig:lifecycle} finalises the arguments in this paper with a vision of three possible life cycles for one specific dataset. The workflow scripts the way they are handle today have one option. The scientist responsible for these scripts tried to optimise the available resources \textbf{manuallly} and intuitively before deliver the final version. Wouldn't be better to have the machine automatising and making such decisions?

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{lifecycle}
  \caption{Life cycle of one dataset: \textbf{C}reate, \textbf{M}igrate, \textbf{R}ead, and \textbf{D}elete}
  \label{fig:lifecycle}
\end{figure}

In this case, we have extra information indicating why Tasks \texttt{ver} and \texttt{post} depend on Task \texttt{model}. In fact, the dependency is not to the complete task, but of some datasets produced by Task \texttt{model}. If this information is available a priori, then we can use it to start Task \texttt{ver} once we have the dataset \texttt{I/O 1} and start Task \texttt{post} once we have the dataset \texttt{I/O 3}, making the general process faster. Still considering this small example, we can assume that dataset \texttt{I/O 2} will not be used for other tasks. So this data can be stored in a persistent storage or it can even be deleted.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{model5}
  \caption{5.}
  \label{fig:model5}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{model6}
  \caption{6.}
  \label{fig:model6}
\end{figure}
